---
title: "Data Cleaning"
author: "Yvonne JIN"
date: "12/18/2020"
output: html_document
---

```{r setup, include=FALSE}
library(foreign)
library(haven)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(lavaan)
library(apaTables)
```

## Loading data

loading database with variable labels

```{r cars}
# read in SPSS data with labels being added to demographic data
df_labelled <- read.spss("merged_Survey on Sleep Health+followedup.sav",use.value.labels = T,to.data.frame = T)



```


Eligibility & consent check
```{r}

parti_lablled <- df_labelled %>% filter(Eli1 == "Yes" & Eli2 == "Yes")
# 1633 -> 1591

parti_lablled <- parti_lablled %>% filter(agree == "Yes")
# 1591 -> 1387


COVID_cleaned <- parti_lablled[c("SCOV_ID")]




```




loading database for numerical computation
```{r}
# read in SPSS data with original (numerical) values
df <- read_spss("merged_Survey on Sleep Health+followedup.sav")

df_1 <- df %>% filter(Eli1 == 2 & Eli2 == 1 & agree == 1)
# 1633 -> 1387

# count No. of NAs for each value
countna <- colSums(is.na(df))
```


## DASS22 psychiatric illnesses


DASS22: Have you ever been diagnosed with mental health problems?


```{r}
#!is.na(parti_lablled$DASS22a_1)

current_medical_illness <- select(parti_lablled, starts_with("DASS22")) %>% 
  filter(!is.na(DASS22))

summary(current_medical_illness)

```


## phyhealth: medical conditions

```{r}

phyhealth_ <- parti_lablled %>% select(starts_with("phyhealth"))

summary(phyhealth_)

table(phyhealth_$phyhealth_10_TEXT)
table(phyhealth_$phyhealth_12_TEXT)

phyhealth_n <- df_1 %>% select(starts_with("phyhealth")) %>%
  mutate(n_medical_illness = phyhealth_1 + phyhealth_2 + phyhealth_3 + phyhealth_4 + phyhealth_5 + phyhealth_6 + phyhealth_7+ phyhealth_8 + phyhealth_9 + phyhealth_10 + phyhealth_11+ phyhealth_12 -12 )

summary(phyhealth_n)

mean(phyhealth_n$n_medical_illness,na.rm = T)

table(phyhealth_n$n_medical_illness)

var(phyhealth_n$n_medical_illness,na.rm = T)

sqrt(0.3180228)

```


## exercise & sedentary hours

exercise

```{r}

# compute active exercising time (in minutes), intensive activity cound as twice of moderate act
df_1 <- df_1 %>% mutate(
  intensive_time =  case_when(strenuous_act1==0 ~ 0, 
                              strenuous_act1a ==1 ~ 20, 
                              strenuous_act1a ==2 ~ 45,
                              strenuous_act1a ==3 ~ 90,
                              strenuous_act1a ==4 ~ 120),
  moderate_time = case_when(moderate_act1 ==0 ~ 0, 
                            moderate_act1a ==1 ~ 20, 
                            moderate_act1a ==2 ~ 45,
                            moderate_act1a ==3 ~ 90,
                            moderate_act1a ==4 ~ 120),
  exercise = as.numeric(strenuous_act1)*intensive_time*2 + as.numeric(moderate_act1)*moderate_time,
  
  intensive_time_F =  case_when(strenuous1_F == 0 ~ 0, 
                                strenuous2_F == 1 ~ 20, 
                                strenuous2_F ==2 ~ 45,
                                strenuous2_F ==3 ~ 90,
                                strenuous2_F ==4 ~ 120),
  moderate_time_F = case_when(moderate1_F == 0 ~ 0, 
                              moderate2_F ==1 ~ 20, 
                              moderate2_F ==2 ~ 45,
                              moderate2_F ==3 ~ 90,
                              moderate2_F ==4 ~ 120),
  exercise_F = as.numeric(strenuous1_F)*intensive_time_F*2 + as.numeric(moderate1_F)*moderate_time_F
)

```

for table 2
```{r}

# Vigorous and moderate activities (VMA)
VMA <- df_1 %>% mutate(VMA_cond = case_when(
  exercise <= 30 ~ "< 30 min",
  exercise > 30 & exercise <= 149 ~ " 30 to 149 min",
  exercise >150 ~ " > 150 min",
  exercise == "NA" ~ "NA"),
                       VMA_F_cond = case_when(
  exercise_F <= 30 ~ "< 30 min",
  exercise_F > 30 & exercise <= 149 ~ " 30 to 149 min",
  exercise_F >150 ~ " > 150 min",
  exercise_F == "NA" ~ "NA")     )

table(VMA$VMA_cond)
table(VMA$VMA_F_cond)


```

sedentary

```{r}


# compute sedentary hours: workday *5 + non-workday *2
df_1 <- df_1 %>% mutate(
  sedentary = sedentary1*5 + sedentary2*2,
  sedentary_F = sedentary1_F*5 + sedentary2_F*2
)

#hist(df_1$sedentary)

summary(df_1$sedentary)
var(df_1$sedentary,na.rm = T)
summary(df_1$sedentary_F)
var(df_1$sedentary_F,na.rm = T)




```

Walkig time 

```{r}


summary(parti_lablled$walking1a)


df_1 <- df_1 %>% mutate(
  walking_average =  case_when(walking1a ==1 ~ 20, 
                            walking1a ==2 ~ 45,
                            walking1a ==3 ~ 90,
                            walking1a ==4 ~ 120),
  
  walking_average_F =  case_when(walk2_F ==1 ~ 20, 
                              walk2_F ==2 ~ 45,
                              walk2_F ==3 ~ 90,
                             walk2_F ==4 ~ 120),
  
  walking_total = as.numeric(walking1)*walking_average,

  walking_total_F = as.numeric(walk1_F)*walking_average_F
)


summary(df_1$walking_total)
var(df_1$walking_total,na.rm = T)
summary(df_1$walking_total_F)
var(df_1$walking_total_F,na.rm = T)
```

write out for Table 3 

```{r}

COVID_cleaned <- COVID_cleaned %>% cbind(exercise = df_1$exercise, 
                                         exercise_F = df_1$exercise_F, 
                                         sedentary = df_1$sedentary, 
                                         sedentary_F = df_1$sedentary_F, 
                                         walking_total = df_1$walking_total,
                                         walking_total_F = df_1$walking_total_F)

```


## COVID impact 
Q2,7,11a,12 - worried/affected, %
Q3 - behavior change, self protection 0/1
Q9 - behavior change, decrease activities, %
Q13 - attitude about government response, %
Q14 - agree with government approaches, %
```{r}
COVID_impact <- df_1 %>% select(starts_with("COVID")) 

COVID_impact <- COVID_impact %>% mutate(worry = (covid2_1+covid2_2+covid7_1+covid11a_1+covid12_1)/5,
                                        decrease_act = (covid9_1+covid9_2+covid9_3+covid9_4+covid9_5)/5,
                                        attituede_ = (covid13_1+covid13_2+covid14_1)/3)

COVID_cleaned <- COVID_cleaned %>% cbind(worry = COVID_impact$worry,
                                         decrease_activity = COVID_impact$decrease_act,
                                         dissatisfaction = COVID_impact$attituede_)

summary(COVID_impact$worry)
summary(COVID_impact$decrease_act)
summary(COVID_impact$attituede_)

COVID_impact$covid14_1

```


## Smoking

```{r}

smoking <- parti_lablled %>% select(starts_with("smoke")) %>% mutate(amount_day = as.integer(smoke1a))


table(smoking$smoke1)

smoke_amount <- smoking %>% filter(smoke1 == "Yes" & !is.na(amount_day))

mean(smoke_amount$amount_day)



```

## drinking 


frequency

```{r}
Drinking <- parti_lablled %>% select(starts_with("alcohol"), "last7days_alcohol_F", "MarApr_alcohol5_F")

# calculate frequency 
Drinking <- Drinking %>% mutate(
  freq_fac =  case_when(alcohol1_2_1 == "Per Year" ~ 1/52, 
                        alcohol1_2_1 == "Per month" ~ 1/4, 
                        alcohol1_2_1 == "Weekly" ~ 1),
  alcohol_freq = (alcohol1_1_1_1)*freq_fac
)

# time 1 
summary(Drinking$alcohol_freq)
var(Drinking$alcohol_freq,na.rm = T)

# March & April
summary(Drinking$alcohol1_F)
var(Drinking$alcohol1_F,na.rm = T)

# last week
summary(Drinking$last7days_alcohol_F)
var(Drinking$last7days_alcohol_F,na.rm = T)



COVID_cleaned <- COVID_cleaned %>% cbind(drinking_freq_time1 = Drinking$alcohol_freq, 
                                         drinking_freq_MarApr = Drinking$alcohol1_F, 
                                         drinking_freq_lastweek =Drinking$last7days_alcohol_F)

```


binge drinking (per week)

```{r}

Binge_drinking <- Drinking %>% mutate(binge_lastyear = as.integer(alcohol2)/52,
                                      binge_mar_apr = as.integer(MarApr_alcohol5_F)/8,
                                      binge_7d = as.integer(alcohol2_F)) 

# time 1 (last year?)

summary(Binge_drinking$binge_lastyear)
var(Binge_drinking$binge_lastyear,na.rm = T)


# Mar & Apr


summary(Binge_drinking$binge_mar_apr)
var(Binge_drinking$binge_mar_apr,na.rm = T)

# last week

summary(Binge_drinking$binge_7d)
var(Binge_drinking$binge_7d,na.rm = T)


COVID_cleaned <- COVID_cleaned %>% cbind(binge_drinking_time1 = Binge_drinking$binge_lastyear, 
                                         binge_drinking_MarApr = Binge_drinking$binge_mar_apr, 
                                         binge_drinking_lastweek =Binge_drinking$binge_7d)

```


## media consumption

```{r}


Media <- df_1 %>% select(starts_with("media"))

summary(Media)


```


## Seeking Help

```{r}

Seek_help <- parti_lablled %>% select(starts_with("seek")) 
  
  
write.csv(Seek_help, "Seek_help_0127.csv",row.names = F, fileEncoding = "UTF-8")

getwd()

```


## additional Qs for sleep problems - graphs


adds1
```{r not used, eval=FALSE, include=FALSE}

add <- parti_lablled %>% select(starts_with("add")) %>% 
                         mutate(adds1_ = case_when(
                           adds1 == "Before 2019" ~ "Before 2019",
                           adds1 == "BetweenJan 01 2019 and March 31 2019" ~ "2019 S1",
                           adds1 == "Between April 01 2019 and June 30 2019" ~ "2019 S2",
                           adds1 == "Between July 01 2019 and September 30 2019" ~ "2019 S3",
                           adds1 == "Between October 01 2019 and Dec 31 2019 " ~ "2019 S4",
                           adds1 == "Between Jan 01 2020 and March 31 2020" ~ "2020 S1",
                           adds1 == "After April 01 2020" ~ "After 2020 S1"))

```


```{r}

insomnia_onset <- data.frame(Time = c("Before 2019","2019 S1", "2019 S2", "2019 S3", "2019 S4", "2020 S1", "After 2020 S1"),
                             value = c(112, 14, 60, 110, 70, 189, 38))

head(insomnia_onset)

ggplot(insomnia_onset, aes(x = "", y=value, fill=Time)) +
  geom_bar(width = 1,stat="identity") +
  coord_polar("y",start=0) +
  fill_palette(palette =  "uchicago") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5),
            size = 4.7) +
  labs(title = "Onset time of Sleep problems",
       y = "" )



```

adds2 & adds3

```{r}

insomnia_cause <- parti_lablled %>% select(starts_with("adds2"), starts_with("adds3")) %>%
  mutate(Cause = case_when(adds2 == "yes" & adds3 == "Yes" ~ "Both",
                           adds2 == "yes" & adds3 == "No" ~ "Protests",
                           adds2 == "no" & adds3 == "Yes" ~ "COVID_19 Outbreak",
                           adds2 == "no" & adds3 == "No" ~ "Neither")
          ) %>% 
  group_by(Cause)  %>%
  count()
         
insomnia_cause <- drop_na(insomnia_cause)


ggplot(insomnia_cause, aes(x = "", y=n, fill=Cause)) +
  geom_bar(width = 1,stat="identity") +
  coord_polar("y",start=0) +
  fill_palette(palette =  "uchicago") +
  geom_text(aes(label = n),
            position = position_stack(vjust = 0.5),
            size = 4.7) +
  labs(title = "Cause of Sleep problems",
       y = "" )


```



adds4 

```{r}

adds4 <- parti_lablled %>% select(starts_with("adds4")) 




  
write.csv(adds4, "adds4_reason_for_sleep_prob_0127.csv",row.names = F, fileEncoding = "UTF-8")




```





## other unused lifestyle data
```{r}


descriptive_ <- df_1 %>% select(starts_with("add"),starts_with("social"),ends_with("impact"))


COVID_cleaned <- COVID_cleaned %>% cbind(descriptive_[1:length(colnames(descriptive_))])

write.csv(COVID_cleaned, "additional_variable_cleaned_0211.csv",row.names = F, fileEncoding = "UTF-8")




```

