---
title: "COVID data correlation analysis"
author: "Yvonne JIN"
date: "2020/9/17"
output: pdf_document
---


This file examine the correlation and change of test scores including ISI, DASS, PSQI based on catagorations of demographic variables.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)

#Libraries
library(tidyverse)
library(tidyr)


#Set Working Directory
setwd("C:/Users/jinya/OneDrive - connect.hku.hk/SLASH Lab/COVID and Sleep data")


# import cleaned data
COVID_cleaned <-read.csv("COVID_cleaned.csv", header = T, fileEncoding = "UTF-8")


```

## ISI

### correlation of ISI and demographic variables

no significant finding.


```{r ISI, echo=TRUE}

corr_text_ISI <- function(x){
  cor.test(COVID_cleaned$ISI_total,x)
}


lapply(COVID_cleaned[,2:9], corr_text_ISI )


corr_text_ISI_F <- function(x){
  cor.test(COVID_cleaned$ISI_F_total,x)
}

lapply(COVID_cleaned[,2:9], corr_text_ISI_F )
# gender and marital closest to threshold


```


### T test for follow up study

no significant finding
```{r}
# T tests
# Is the difference in means across two groups statistically significant?


# overall, p-value = 0.05116
t.test(COVID_cleaned$ISI_total, COVID_cleaned$ISI_F_total)

#paired t.test()!!


#--------------------
  
# female, p-value = 0.06312, mean 10.704225 -> 9.482993 
t.test(COVID_cleaned[COVID_cleaned$Gender == 2,"ISI_total"], COVID_cleaned[COVID_cleaned$Gender == 2,'ISI_F_total'])
# male, p-value = 0.4268, mean 11.30769 -> 10.71429 
t.test(COVID_cleaned[COVID_cleaned$Gender == 1,"ISI_total"], COVID_cleaned[COVID_cleaned$Gender == 1,'ISI_F_total'])

t.test(COVID_cleaned[COVID_cleaned$Gender == 0,"ISI_total"], COVID_cleaned[COVID_cleaned$Gender == 1,'ISI_total'])


#--------------

# Marital

t_test_marital <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Marital == x,"ISI_total"], COVID_cleaned[COVID_cleaned$Marital == x,'ISI_F_total'])
}


t_test_marital(1) 
t_test_marital(2)
t_test_marital(3)
t_test_marital(4)
t_test_marital(5)

# -------------------


# education
#no significant finding
t_test_education <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Education == x,"ISI_total"], COVID_cleaned[COVID_cleaned$Education == x,'ISI_F_total'])
}

hist(COVID_cleaned$Education)

t_test_education(2)
t_test_education(3)
t_test_education(4) 
t_test_education(5)
t_test_education(7)

# age
# no significant findings
t_test_age <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Age == x,"ISI_total"], COVID_cleaned[COVID_cleaned$Age == x,'ISI_F_total'])
}

hist(COVID_cleaned$Age)

t_test_age(1)
t_test_age(2)
t_test_age(3)
t_test_age(4)
t_test_age(5)
t_test_age(7)

# ......................................................to be continued

```
## DASS

### correlation of DASS and demographic variables

Age: Young ppl are more likely to be anxious, cor = -0.15, p-value = 0.01431


```{r DASS, echo=TRUE}

corr_text <- function(x){
  cor.test(COVID_cleaned$DASS_total,x)
}

lapply(COVID_cleaned[,2:9], corr_text )


corr_text_F <- function(x){
  cor.test(COVID_cleaned$DASS_F_total,x)
}

lapply(COVID_cleaned[,2:9], corr_text_F )
# gender p-value = 0.05049 cor = -0.1223759 , age p-value = 0.07211, p-value = 0.1013

```
Marital: 

1 - single
2 - in a relationship
3 - cohabitation
4 - Married
5 - devorced? (very few observations)
6 - widowed
ppl not in a stable relationship have higher score?
cor = -0.1830635,p-value = 0.003287

excluding cohabitation and devorced individual:
cor = -0.1816677 , p-value = 0.004176

```{r DASS Marital }

# explore marriage data
plot(COVID_cleaned$Marital, COVID_cleaned$DASS_total)
hist(COVID_cleaned$Marital)
df_married <- select(COVID_cleaned, "DASS_total", "Marital")
df_married <- df_married[df_married$Marital != 3 & df_married$Marital != 5,]
cor.test(df_married$DASS_total, df_married$Marital)


```


### T test for follow up study

the symptoms of female reduced significantly 
```{r}
# T tests
# Is the difference in means across two groups statistically significant?


# overall, p-value = 0.05016
t.test(COVID_cleaned$DASS_total, COVID_cleaned$DASS_F_total)

# Gender

# female, p-value = 0.04938, mean 44.308 -> 38.510
t.test(COVID_cleaned[COVID_cleaned$Gender == 2,"DASS_total"], COVID_cleaned[COVID_cleaned$Gender == 2,'DASS_F_total'])
# male, p-value = 0.4372, mean 48.13084 -> 45.12150 
t.test(COVID_cleaned[COVID_cleaned$Gender == 1,"DASS_total"], COVID_cleaned[COVID_cleaned$Gender == 1,'DASS_F_total'])

## female more significant, p-values very similar for ISI and DASS scores

#------------

# Marital

t_test_marital <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Marital == x,"DASS_total"], COVID_cleaned[COVID_cleaned$Marital == x,'DASS_F_total'])
}


t_test_marital(1) # p-value = 0.04989, mean  50.33083 -> 43.86466 
t_test_marital(2)
t_test_marital(3)
t_test_marital(4)
t_test_marital(5)


#---------------

# education

t_test_education <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Education == x,"DASS_total"], COVID_cleaned[COVID_cleaned$Education == x,'DASS_F_total'])
}

hist(COVID_cleaned$Education)

t_test_education(2)
t_test_education(3)
t_test_education(4) # p-value = 0.04759
t_test_education(5)
t_test_education(7)


#-------------------
  
# age
# no significant findings
t_test_age <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Age == x,"DASS_total"], COVID_cleaned[COVID_cleaned$Age == x,'DASS_F_total'])
}

hist(COVID_cleaned$Age)

t_test_age(1)
t_test_age(2)
t_test_age(3)
t_test_age(4)
t_test_age(5)
t_test_age(7)

# ......................................................to be continued

```
## PSQI


### correlation 

all not significant.


```{r ISI, echo=TRUE}

corr_text_PSQI <- function(x){
  cor.test(COVID_cleaned$PSQI_total,x)
}

lapply(COVID_cleaned[,2:9], corr_text_PSQI )


corr_text_PSQI_F <- function(x){
  cor.test(COVID_cleaned$PSQI_F_total,x)
}


lapply(COVID_cleaned[,2:9], corr_text_PSQI_F )


```


### T test for follow up study

the symptoms of female reduced significantly 
```{r}
# T tests
# Is the difference in means across two groups statistically significant?


# overall,  p-value = 0.112, not significant
t.test(COVID_cleaned$PSQI_total, COVID_cleaned$PSQI_F_total)


#--------------------
  
# female, p-value = 0.03401, mean  8.590604 -> 7.791946 
t.test(COVID_cleaned[COVID_cleaned$Gender == 2,"PSQI_total"], COVID_cleaned[COVID_cleaned$Gender == 2,'PSQI_F_total'])
# male, p-value = 1, no effect at all?
t.test(COVID_cleaned[COVID_cleaned$Gender == 1,"PSQI_total"], COVID_cleaned[COVID_cleaned$Gender == 1,'PSQI_F_total'])


#--------------

# Marital

t_test_marital <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Marital == x,"PSQI_total"], COVID_cleaned[COVID_cleaned$Marital == x,'PSQI_F_total'])
}


t_test_marital(1) 
t_test_marital(2)
t_test_marital(3)
t_test_marital(4) # p-value = 1
t_test_marital(5) # p-value = 1

# -------------------


# education
#no significant finding
t_test_education <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Education == x,"PSQI_total"], COVID_cleaned[COVID_cleaned$Education == x,'PSQI_F_total'])
}

hist(COVID_cleaned$Education)

t_test_education(2)
t_test_education(3)
t_test_education(4) 
t_test_education(5)
t_test_education(7)

# age
# no significant findings
t_test_age <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Age == x,"PSQI_total"], COVID_cleaned[COVID_cleaned$Age == x,'PSQI_F_total'])
}

hist(COVID_cleaned$Age)

t_test_age(1)
t_test_age(2) #p-value = 0.07319
t_test_age(3)
t_test_age(4)
t_test_age(5)
t_test_age(7)

# ......................................................to be continued

```

## QoL


Marital: p-value = 0.09404, similar trend?   cor = 0.1055008

Occupation_changed: p-value = 0.01725, cor = -0.149607 ?

Occupation:  p-value = 0.2084

1 - student
2 - self-employed
3 - unemployed
4 - homemaker
5 - retired
autonomous/stable -> better QoL?

---------
Follow up

Marital:  p-value = 0.0334, cor = 0.1332714 

1 - single
2 - in a relationship
3 - cohabitation
4 - Married
5 - devorced? (very few observations)
6 - widowed

```{r QoL}

corr_text_QoL <- function(x){
  cor.test(COVID_cleaned$QoL_score,x)
}

lapply(COVID_cleaned[,2:9], corr_text_QoL )


corr_text_QoL_F <- function(x){
  cor.test(COVID_cleaned$QoL_F_score,x)
}


lapply(COVID_cleaned[,2:9], corr_text_QoL_F )

```

excluding cohabitation and divorced: p-value = 0.04266, cor = 0.1298845 

```{r QoL Marital }

# explore marriage data
plot(COVID_cleaned$Marital, COVID_cleaned$QoL_score)

df_married <- select(COVID_cleaned, "QoL_score", "Marital")
df_married <- df_married[df_married$Marital != 3 & df_married$Marital != 5,]
cor.test(df_married$QoL_score, df_married$Marital)


```



### T test for follow up study


```{r}
# T tests
# Is the difference in means across two groups statistically significant?


# overall,  p-value = 0.1005, not significant
t.test(COVID_cleaned$QoL_score, COVID_cleaned$QoL_F_score)


#--------------------
  
# female,p-value = 0.07287, mean   16.15646 ->  17.50336  
t.test(COVID_cleaned[COVID_cleaned$Gender == 2,"QoL_score"], COVID_cleaned[COVID_cleaned$Gender == 2,'QoL_F_score'])
# male, p-value = 0.6589
t.test(COVID_cleaned[COVID_cleaned$Gender == 1,"QoL_score"], COVID_cleaned[COVID_cleaned$Gender == 1,'QoL_F_score'])


#--------------

# Marital
# no significant finding
t_test_marital <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Marital == x,"QoL_score"], COVID_cleaned[COVID_cleaned$Marital == x,'QoL_F_score'])
}


t_test_marital(1) 
t_test_marital(2)
t_test_marital(3)
t_test_marital(4) 
t_test_marital(5) 

# -------------------


# education
#no significant finding
t_test_education <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Education == x,"QoL_score"], COVID_cleaned[COVID_cleaned$Education == x,'QoL_F_score'])
}

hist(COVID_cleaned$Education)

t_test_education(2)
t_test_education(3)
t_test_education(4) 
t_test_education(5)
t_test_education(7)

# age
# no significant findings
t_test_age <- function(x){
  t.test(COVID_cleaned[COVID_cleaned$Age == x,"QoL_score"], COVID_cleaned[COVID_cleaned$Age == x,'QoL_F_score'])
}

hist(COVID_cleaned$Age)

t_test_age(1)
t_test_age(2)
t_test_age(3)
t_test_age(4)
t_test_age(5)
t_test_age(7)

# ......................................................to be continued

```
```{r}
save.image("correlation.Rdata")
```


