---
title: "COVID_Sleep_survey_data_cleaning"
author: "Yvonne JIN"
date: "2020/9/8"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data


```{r prep, message=FALSE, warning=FALSE, include=FALSE}
#Libraries
library(haven)
library(tidyverse)
library(tidyr)
library(ggplot2)


#Set Working Directory
setwd("C:/Users/jinya/OneDriveq - connect.hku.hk/SLASH Lab/COVID and Sleep data")


#Reading source files
df <- read_spss("merged_Survey on Sleep Health+followedup.sav")

#write.csv(df,"merged_survey.csv",fileEncoding = UTF-8)
#colnames(df)
```

Filter out non-complete individual

```{r finish, echo=TRUE, message=TRUE, warning=FALSE}

df %>% group_by(Finished) %>%
  count()
# 986 finished out of 1633

df_1 <- filter(df,Finished == 1) 

## add Demographic data to cleaned data frame
COVID_cleaned <- df_1[c("SCOV_ID", "Gender", "Age", "Ethnicity", "Marital", "Education", "Occupation", "income", "household_size")]

#change coding value of occupation: 1, 32, 33, 34, 35 -> 1,2,3,4,5
#COVID_cleaned <- mutate(COVID_cleaned,Occupation = ifelse(Occupation == 32, 2, ifelse(Occupation == 33, 3, ifelse(Occupation == 34, 4,ifelse(Occupation == 35, 5, 1)))))

# HK residence : 2 -> yes, 3 -> no
COVID_cleaned$HK_residence <- ifelse(df_1$Eli1 == 2, T, F)

COVID_cleaned$stayed_InHK <- ifelse(df_1$Eli2 == 1, T, F)

```



## comuting PSQI Score
```{r PSQI, echo=TRUE, warning=FALSE}

#Calculate PSQI scores
PSQI_temp <- select(df_1, SCOV_ID, starts_with("PSQI")& !ends_with("_F"))

#Compute C1: Subjective sleep quality
# Q11 in 1st survey, Q12 in follow-up survey
PSQI_temp <- mutate(PSQI_temp, C1 = PSQI11)


#compute C2: Sleep latency
#Compute C2a: 
#time needed to fall asleep (DIS)
# if fill in hours but not minutes, taken as 0
PSQI_temp <- mutate(PSQI_temp, 
                    PSQI4_1_1a = ifelse(is.na(PSQI4_1_1), 0, PSQI4_1_1),
                    PSQI4_2_1a = ifelse(is.na(PSQI4_2_1), 0, PSQI4_2_1),
                    DIS = PSQI4_1_1a * 60 + PSQI4_2_1a)
PSQI_temp <- mutate(PSQI_temp, C2a = case_when(DIS <= 15 ~ 0, DIS %in% 16:30 ~ 1, DIS %in% 31:60 ~ 2, DIS > 60 ~ 3))
#Compute C2b: DIS + Q5a (Q6_1) trouble fall asleep within 30min
PSQI_temp <- mutate(PSQI_temp, C2b = C2a + PSQI6_1)
#Compute C2
PSQI_temp <- mutate(PSQI_temp, C2 = case_when(C2b == 0 ~ 0, C2b %in% 1:2 ~ 1, C2b %in% 3:4 ~ 2, C2b > 4 ~ 3))



#Compute C3: Sleep duration
#Compute sleep duration (SD)
# if specified hours but not minutes, taken PSQI5_2_1 as 0
PSQI_temp <- mutate(PSQI_temp, SD = ifelse(!is.na(PSQI5_2_1),PSQI5_1_1 + PSQI5_2_1/60,PSQI5_1_1))
PSQI_temp <- mutate(PSQI_temp, C3 = ifelse(SD >= 7, 0,
                                    ifelse(SD >= 6, 1,
                                    ifelse(SD >= 5, 2, 3))))

#Compute C4:  Sleep efficiency
#  Sleep efficiency = (# hours slept/# hours in bed) X 100%  //hours? I used minutes

#Compute time in bed
# mutate NA->0 for minutes (PSQI3_2_1, PSQI_2_2_1)
PSQI_temp <- mutate(PSQI_temp, PSQI2_2_1a = ifelse(is.na(PSQI2_2_1),0,PSQI2_2_1),
                               PSQI3_2_1a = ifelse(is.na(PSQI3_2_1),0,PSQI3_2_1))
PSQI_temp <- mutate(PSQI_temp, bedtime_a = (PSQI3_1_1*60 + PSQI3_2_1a - PSQI2_1_1*60 - PSQI2_2_1a),
                               bedtime = ifelse(bedtime_a>0, bedtime_a, (24*60+bedtime_a))) # add one day if negative # don't use hours: floor division lost decimal.
                              
#Compute sleep duration in minutes
# if specified hours but not minutes, taken PSQI5_2_1 as 0
PSQI_temp <- mutate(PSQI_temp, sleeptime = ifelse(!is.na(PSQI5_2_1),PSQI5_1_1*60 + PSQI5_2_1,PSQI5_1_1*60))

#Compute SE
PSQI_temp <- mutate(PSQI_temp, SE = sleeptime/bedtime)
PSQI_temp <- mutate(PSQI_temp, C4 = ifelse(SE >= 0.85, 0,
                                    ifelse(SE >= 0.75, 1,
                                    ifelse(SE >= 0.65, 2, 3))))



#Compute C5:  Sleep disturbance
#Compute accumulative scores
PSQI_temp <- mutate(PSQI_temp, C5a = PSQI6_2+PSQI6_3+PSQI6_4+PSQI6_5+PSQI6_6+PSQI6_7+PSQI6_8+PSQI6_9)
PSQI_temp <- mutate(PSQI_temp, C5b = ifelse(PSQI7 == 1, C5a + PSQI7b, C5a)) # other causes

PSQI_temp <- mutate(PSQI_temp, C5 = ifelse(C5b == 0, 0,
                                    ifelse(C5b <= 9, 1,
                                    ifelse(C5b <= 18, 2, 3))))


#Compute C6: Use of sleep medication
PSQI_temp <- mutate(PSQI_temp, C6 = ifelse(is.na(PSQI8),0,PSQI8))



#Compute C7: Daytime dysfunction
#Compute C7a
PSQI_temp <- mutate(PSQI_temp, 
                    PSQI9a = ifelse(is.na(PSQI9),0,PSQI9),
                    PSQI10a = ifelse(is.na(PSQI10),0,PSQI10),
                    C7a = PSQI9a + PSQI10a)
PSQI_temp <- mutate(PSQI_temp, C7 = ifelse(C7a %in% 1:2, 1,
                                    ifelse(C7a %in% 3:4, 2, 3)))
#Compute PSQI composite score for each participant
PSQI_temp <- mutate(PSQI_temp, PSQI_total = C1 + C2 + C3 + C4 + C5 + C6 + C7, sleep_duration = sleeptime)  
PSQI_score <- PSQI_temp %>% drop_na(PSQI_total,sleep_duration)

#glimpse(PSQI_score)


#install.packages("psy")
library(psy)

PSQI_alpha <- select(PSQI_temp, C1, C2, C3, C4, C5, C6, C7, PSQI_total)

cronbach(PSQI_alpha)


```

```{r}

# export data 
PSQI_score <- left_join(PSQI_score, df, by = "SCOV_ID") # 710 individual # 838 after manually checking for NA
write.csv(PSQI_score, "PSQI_unfiltered_score.csv")

PSQI_na <- filter(PSQI_temp, is.na(PSQI_total)) #795
write.csv(PSQI_na,"PSQI_unfiltered_na.csv")


```

```{r}
# export data_filterd "Finish ==1"
write.csv(PSQI_score, "PSQI_filtered_score.csv") # 599 individual # 896 after manually checking for NA


# export NA
PSQI_na <- filter(PSQI_temp, is.na(PSQI_total)) # 388 -> 90
write.csv(PSQI_na,"PSQI_na_checked.csv")


```



follow up survey

```{r echo=FALSE}

#Calculate PSQI scores
PSQI_F_temp <- select(df_1, SCOV_ID, starts_with("PSQI") & ends_with("_F"))

#Compute C1: Subjective sleep quality
# Q11 in 1st survey, Q12 in follow-up survey
PSQI_F_temp <- PSQI_F_temp %>% 
  mutate(PSQI1_2_1_Fa = ifelse(is.na(PSQI1_2_1_F),0,PSQI1_2_1_F),# go to sleep
         PSQI2_2_1_Fa = ifelse(is.na(PSQI2_2_1_F),0,PSQI2_2_1_F),# get up
         PSQI3_1_1_Fa = ifelse(is.na(PSQI3_1_1_F),0,PSQI3_1_1_F),
         PSQI3_2_1_Fa = ifelse(is.na(PSQI3_2_1_F),0,PSQI3_2_1_F), # falling asleep
         PSQI4_2_1_Fa = ifelse(is.na(PSQI4_2_1_F),0,PSQI4_2_1_F), #sleep duration
         PSQI9_Fa = ifelse(is.na(PSQI9_F),0,PSQI9_F), # sleep medicine
         PSQI10_Fa = ifelse(is.na(PSQI10_F),0,PSQI10_F),
         PSQI11_Fa = ifelse(is.na(PSQI11_F),0,PSQI11_F), # daytime disfunction
         
        
         C1 = PSQI12_F, # subjective quality
         DIS = PSQI3_1_1_F * 60 + PSQI3_2_1_Fa, 
         C2a = case_when(DIS <= 15 ~ 0, DIS %in% 16:30 ~ 1, DIS %in% 31:60 ~ 2, DIS > 60 ~ 3),
         C2b = C2a + PSQI5_1_F,
         C2 = case_when(C2b == 0 ~ 0, C2b %in% 1:2 ~ 1, C2b %in% 3:4 ~ 2, C2b > 4 ~ 3),
         SD = PSQI4_1_1_F + PSQI4_2_1_Fa/60, # sleep duration
         C3 = ifelse(SD >= 7, 0,
                                    ifelse(SD >= 6, 1,
                                    ifelse(SD >= 5, 2, 3))),
         bedtime_a = (PSQI2_1_1_F*60 + PSQI2_2_1_Fa - PSQI1_1_1_F*60 - PSQI1_2_1_Fa),
         bedtime = ifelse(bedtime_a>0, bedtime_a, (24*60+bedtime_a)),
         sleeptime = PSQI4_1_1_F*60 + PSQI4_2_1_Fa,
         SE = sleeptime/bedtime,
         C4 = ifelse(SE >= 0.85, 0,
                     ifelse(SE >= 0.75, 1,
                     ifelse(SE >= 0.65, 2, 3))),
         C5a = PSQI5_2_F+PSQI5_3_F+PSQI5_4_F+PSQI5_5_F+PSQI5_6_F+PSQI5_7_F+PSQI5_8_F+PSQI5_9_F,
         C5b = ifelse(PSQI6_F == 1, C5a + PSQI8_F, C5a), # other causes
         C5 = ifelse(C5b == 0, 0,
                     ifelse(C5b <= 9, 1,
                     ifelse(C5b <= 18, 2, 3))),
         C6 = PSQI9_Fa,
         C7a = PSQI10_Fa + PSQI11_Fa,
         C7 = ifelse(C7a %in% 1:2, 1,
                                    ifelse(C7a %in% 3:4, 2, 3)))
         
         
         

#Compute PSQI composite score for each participant
PSQI_F_temp <- mutate(PSQI_F_temp, PSQI_total = C1 + C2 + C3 + C4 + C5 + C6 + C7, sleep_duration = sleeptime) 
PSQI_F_score <- PSQI_F_temp %>% drop_na(PSQI_total,sleep_duration)

#glimpse(PSQI_F)


PSQI_F_alpha <- select(PSQI_F_temp, C1, C2, C3, C4, C5, C6, C7, PSQI_total)

cronbach(PSQI_F_alpha)


```


```{r}

# export data # unfiltered
PSQI_F_score <- left_join(PSQI_F_score, df, by = "SCOV_ID") # 357 individual
#write.csv(PSQI_F_score, "PSQI_F_unfiltered_score.csv")
#NA
PSQI_F_na <- filter(PSQI_F_temp, is.na(PSQI_total)) # 1276
#write.csv(PSQI_na,"PSQI_unfiltered_na.csv") # all additonal are non-finished individuals. same.


```

```{r}
# export data_filterd "Finish ==1"
write.csv(PSQI_F_score, "PSQI_F_filtered_score.csv") # 357 individual -> 463


# export NA
PSQI_F_na <- filter(PSQI_F_temp, is.na(PSQI_total)) # 629 -> 523
write.csv(PSQI_F_na,"PSQI_F_na.csv")


```


```{r join dataframes, echo=TRUE}

#PSQI_compare <- demographic %>% 
#  left_join(PSQI_score, by = "SCOV_ID") %>%
#  left_join(PSQI_F,by = "SCOV_ID")


# Add score to cleaned data frame
COVID_cleaned <- COVID_cleaned %>% cbind(PSQI_total = PSQI_temp$PSQI_total, Sleep_duration = PSQI_temp$sleep_duration, 
                                         PSQI_F_total = PSQI_F_temp$PSQI_total, Sleep_F_duration =PSQI_F_temp$sleep_duration)


```


```{r for viz, include=FALSE}

demographic <- df_1[c("SCOV_ID", "Gender", "Age", "Ethnicity", "Marital", "Education", "Occupation", "income", "household_size")]


# add columns
PSQI_0 <- demographic %>%
  cbind(PSQI_total = PSQI_temp$PSQI_total, Sleep_duration = PSQI_temp$sleep_duration) 
PSQI_0$status <- 0

PSQI_1 <- demographic %>%
  cbind(PSQI_total = PSQI_F_temp$PSQI_total, Sleep_duration =PSQI_F_temp$sleep_duration)
PSQI_1$status <- 1

PSQI_m <- left_join(PSQI_0,PSQI_1,by = "SCOV_ID") %>% drop_na(PSQI_total.y)
PSQI_m$change <- PSQI_m$PSQI_total.y - PSQI_m$PSQI_total.x


PSQI_compare <- rbind(PSQI_0,PSQI_1)


PSQI_compare <- PSQI_compare %>% drop_na(PSQI_total)
  
            
```


Plot PSQI comparison 
```{r PSQI compasion, echo=FALSE}

ggplot(PSQI_m) +
  aes(x = change) +
  geom_histogram(binwidth = 1)


ggplot(PSQI_m) +
  aes(x = change, fill = factor(Gender.x)) +
  geom_histogram(binwidth = 1) +
  scale_fill_brewer(palette = "Blues")

ggplot(PSQI_m) +
  aes(x = change, fill = factor(Education.x)) +
  geom_histogram(binwidth = 1) +
  scale_fill_brewer(palette = "Blues")

ggplot(PSQI_m) +
  aes(x = change, fill = factor(income.x)) +
  geom_histogram(binwidth = 1) +
  scale_fill_brewer(palette = "RdBu")


ggplot(PSQI_m) +
  aes(x = change, fill = factor(Marital.x)) +
  geom_histogram(binwidth = 1) +
  scale_fill_brewer(palette = "Blues")
  
```

```{r echo=FALSE}

ggplot(PSQI_compare) +
  aes(x = factor(Age), y = PSQI_total, color = factor(status))  +
  geom_boxplot()

# Ethinicity?
ggplot(PSQI_compare) +
  aes(x = factor(Ethnicity), y = PSQI_total, color =  factor(status))  +
  geom_boxplot()

# Marital status 
ggplot(PSQI_compare) +
  aes(x = factor(Marital), y = PSQI_total, color =  factor(status))  +
  geom_boxplot()

# Education 
ggplot(PSQI_compare) +
  aes(x = factor(Education), y = PSQI_total, color =  factor(status))  +
  geom_boxplot()

# Income
ggplot(PSQI_compare) +
  aes(x = factor(income), y = PSQI_total, color =  factor(status))  +
  geom_boxplot()

# household size
ggplot(PSQI_compare) +
  aes(x = factor(household_size), y = PSQI_total, color =  factor(status))  +
  geom_boxplot()


```



## DASS


```{r DASS, echo=TRUE}

###before---

#Calculate DASS_dep scores for each participant and filter out NAs
df_1 <- mutate(df_1, DASS_dep_total = 2*(DASS_3+DASS_5+DASS_10+DASS_13+DASS_16+DASS_17+DASS_21))

#Calculate DASS_anx scores for each participant and filter out NAs
df_1 <- mutate(df_1, DASS_anx_total = 2*(DASS_2+DASS_4+DASS_7+DASS_9+DASS_15+DASS_19+DASS_20))

#Calculate DASS_str scores for each participant and filter out NAs
df_1 <- mutate(df_1, DASS_str_total = 2*(DASS_1+DASS_6+DASS_8+DASS_11+DASS_12+DASS_14+DASS_18))

# compute DASS total score, max = 21*3*2 = 126
df_1 <- mutate(df_1, DASS_total = DASS_dep_total + DASS_anx_total + DASS_str_total)

df_1 %>% ggplot() +
  aes(x = DASS_total) +
  geom_histogram(bins = 100) +
  labs(title = "DASS Total score")
  
DASS_alpha <- select(df_1, starts_with("DASS_") & !ends_with("_F") & !ends_with("total"))

DASS_alpha <- cbind(DASS_alpha, df_1$DASS_total)

cronbach(DASS_alpha)

#summary(df_1$DASS_total)

```



```{r}
# export data_filtered "Finish ==1"
write.csv(df_1, "DASS_filtered_score.csv") # individual


# export NA
DASS_na <- filter(df_1, is.na(DASS_total))
write.csv(DASS_na,"DASS_na.csv")


```



Follow up survey

```{r DASS_F, echo=FALSE}
####Follow up -----

#Calculate DASS_dep scores for each participant and filter out NAs
df_1 <- mutate(df_1, DASS_dep_F_total = 2*(DASS_3_F+DASS_5_F+DASS_10_F+DASS_13_F+DASS_16_F+DASS_17_F+DASS_21_F))

#Calculate DASS_anx scores for each participant and filter out NAs
df_1 <- mutate(df_1, DASS_anx_F_total = 2*(DASS_2_F+DASS_4_F+DASS_7_F+DASS_9_F+DASS_15_F+DASS_19_F+DASS_20_F))

#Calculate DASS_str scores for each participant and filter out NAs
df_1 <- mutate(df_1, DASS_str_F_total = 2*(DASS_1_F+DASS_6_F+DASS_8_F+DASS_11_F+DASS_12_F+DASS_14_F+DASS_18_F))

## DASS 22_a?

# compute DASS total score
df_1 <- mutate(df_1, DASS_F_total = DASS_dep_F_total + DASS_anx_F_total + DASS_str_F_total)

#na.exclude(df_1)

df_1 %>% ggplot() +
  aes(x = DASS_F_total) +
  geom_histogram(bins = 100) +
  labs(title = "DASS Total score_Follow up")

summary(df_1$DASS_F_total)

####

DASS_F_alpha <- select(df_1, starts_with("DASS_") & ends_with("_F") & !ends_with("total"))

DASS_F_alpha <- cbind(DASS_F_alpha, df_1$DASS_F_total)

cronbach(DASS_F_alpha)


# 503 NA!
```


```{r}

# export data 
DASS_score <- left_join(PSQI_score, df, by = "SCOV_ID") # 710 individual
write.csv(PSQI_score, "PSQI_unfiltered_score.csv")

DASS_na <- filter(df_1, is.na(DASS_total))
write.csv(PSQI_na,"DASS_unfiltered_na.csv")


```

```{r}
# export data_filterd "Finish ==1"
write.csv(df_1, "DASS_F_filtered_score.csv") # individual


# export NA
DASS_F_na <- filter(df_1, is.na(DASS_F_total))
write.csv(DASS_F_na,"DASS_F_na.csv")


```


```{r DASS_add, echo=TRUE}

# addd to cleaned data frame

COVID_cleaned <- COVID_cleaned %>% cbind(DASS_total = df_1$DASS_total, 
                                         DASS_F_total = df_1$DASS_F_total)


```



Merge the two data set and compare
```{r echo=FALSE}

### Compare ###
# in a different data frame

DASS_before <- df_1[c("SCOV_ID", "Gender", "Age", "Ethnicity", "Marital", "Education", "Occupation", "income", "household_size","DASS_total")]

colnames(DASS_before) = c("SCOV_ID", "Gender", "Age", "Ethnicity", "Marital", "Education", "Occupation", "income", "household_size","DASS_score")

DASS_before$status <- 0

#rename(DASS_before, DASS_score = DASS_total) ?



DASS_after <- df_1[c("SCOV_ID", "Gender", "Age", "Ethnicity", "Marital", "Education", "Occupation", "income", "household_size",DASS_Score = "DASS_F_total")]

colnames(DASS_after) = c("SCOV_ID", "Gender", "Age", "Ethnicity", "Marital", "Education", "Occupation", "income", "household_size",DASS_Score = "DASS_score")

DASS_after$status <- 1

#names(DASS_after[10])[1] <- "DASS_Score"
#rename(DASS_after, DASS_score = DASS_F_total)

DASS_compare <- rbind(DASS_before, DASS_after) %>% drop_na


ggplot(DASS_compare) +
  aes(x = factor(status), y = DASS_score)  +
  geom_boxplot()


# Age and DASS
ggplot(DASS_compare) +
  aes(x = factor(Age), y = DASS_score, color =  factor(status))  +
  geom_boxplot()

# Ethinicity?
ggplot(DASS_compare) +
  aes(x = factor(Ethnicity), y = DASS_score, color =  factor(status))  +
  geom_boxplot()

# Marital status 
ggplot(DASS_compare) +
  aes(x = factor(Marital), y = DASS_score, color =  factor(status))  +
  geom_boxplot()

# Education 
ggplot(DASS_compare) +
  aes(x = factor(Education), y = DASS_score, color =  factor(status))  +
  geom_boxplot()

# Income
ggplot(DASS_compare) +
  aes(x = factor(income), y = DASS_score, color =  factor(status))  +
  geom_boxplot()

# household size
ggplot(DASS_compare) +
  aes(x = factor(household_size), y = DASS_score, color =  factor(status))  +
  geom_boxplot()
```

## ISI score
```{r}
df_1 <- df_1 %>%mutate(ISI_total = ISI1_1 + ISI1_2 + ISI1_3 + ISI2_1 + ISI3_1 + ISI4_1 + ISI5_1)

df_1 <- df_1 %>%mutate(ISI_F_total = ISI_1_1_F + ISI_1_2_F + ISI_1_3_F + ISI_2_1_F + ISI_3_1_F + ISI_4_1_F + ISI_5_1_F)

COVID_cleaned <- COVID_cleaned %>% cbind(ISI_total = df_1$ISI_total)

COVID_cleaned <- COVID_cleaned %>% cbind(ISI_F_total = df_1$ISI_F_total)

hist(df_1$ISI_total, breaks = 30)

hist(df_1$ISI_F_total, breaks = 30)


ISI_alpha <- select(df_1, starts_with("ISI") & !ends_with("_F"))

ISI_F_alpha <- select(df_1, starts_with("ISI") & ends_with("_F"))

ISI_alpha <-sapply(ISI_alpha, as.integer)

ISI_F_alpha <-sapply(ISI_F_alpha, as.integer)


ISI_compare <- select(df_1,ISI, ISI_total)

ISI_F_compare <- select(df_1,ISI_F, ISI_F_total)

cronbach(ISI_alpha)

cronbach(ISI_F_alpha)


```


## Quality of Life 
```{r}

df_1 <- df_1 %>%mutate(QoL_score = QoL_1 + QoL_2 + QoL_3 + QoL_4 + QoL_5)

df_1 <- df_1 %>%mutate(QoL_F_score = QoL1_1_F + QoL1_2_F +  QoL1_3_F + QoL1_4_F + QoL1_5_F)

COVID_cleaned <- COVID_cleaned %>% cbind(QoL_score = df_1$QoL_score)

COVID_cleaned <- COVID_cleaned %>% cbind(QoL_F_score = df_1$QoL_F_score)

hist(df_1$QoL_score, breaks = 35)

hist(df_1$QoL_F_score, breaks = 35)

```



## Height and Weight
```{r} 

# height: convert inch to cm
df_1 <- df_1 %>% mutate(Height = ifelse(height_2_1 == 1, height_1_1_1, height_1_1_1 * 2.54)) 

# add to cleaned data 
COVID_cleaned <- COVID_cleaned %>% cbind(Height = df_1$Height)

# weight: convert pound to kg
df_1 <- df_1 %>% mutate(Weight = ifelse(weight_2_1 == 2, weight_1_1_1, weight_1_1_1 * 0.45359237)) 

COVID_cleaned <- COVID_cleaned %>% cbind(Weight = df_1$Weight)

# Weight_F:follow up study
df_1 <- df_1 %>% mutate(Weight_F = ifelse(weight_2_1_F == 2, weight_1_1_1_F, weight_1_1_1_F * 0.45359237)) 

COVID_cleaned <- COVID_cleaned %>% cbind(Weight_F = df_1$Weight_F)


```



## physical activity level
#do it seperately


```{r physical activity}

#COVID_cleaned <- COVID_cleaned %>% cbind(strenuous_act1 = df_1$strenuous_act1, strenuous_act1a = df_1$strenuous_act1a, moderate_act1 = df_1$moderate_act1, moderate_act1a = df_1$moderate_act1a, walking1 = df_1$walking1 )


lapply(na.exclude(df_1$strenuous_act1a), duration_conversion)

# compute active exercising time (in minutes), intensive activity cound as twice of moderate act
df_1 <- df_1 %>% mutate(
  intensive_time =  case_when(strenuous_act1a ==1 ~ 20, strenuous_act1a ==2 ~ 45,strenuous_act1a ==3 ~ 90,strenuous_act1a ==4 ~ 120),
  moderate_time = case_when(moderate_act1a ==1 ~ 20, moderate_act1a ==2 ~ 45,moderate_act1a ==3 ~ 90,moderate_act1a ==4 ~ 120),
  exercise = as.numeric(strenuous_act1)*intensive_time*2 + as.numeric(moderate_act1)*moderate_time,
  
  intensive_time_F =  case_when(strenuous2_F ==1 ~ 20, strenuous2_F ==2 ~ 45,strenuous2_F ==3 ~ 90,strenuous2_F ==4 ~ 120),
  moderate_time_F = case_when(moderate2_F ==1 ~ 20, moderate2_F ==2 ~ 45,moderate2_F ==3 ~ 90,moderate2_F ==4 ~ 120),
  exercise_F = as.numeric(strenuous1_F)*intensive_time_F*2 + as.numeric(moderate1_F)*moderate_time_F
  )

# compute sedentary hours: workday *5 + non-workday *2
df_1 <- df_1 %>% mutate(
  sedentary = sedentary1*5 + sedentary2*2,
  sedentary_F = sedentary1_F*5 + sedentary2_F*2
  )
 

COVID_cleaned <- COVID_cleaned %>% cbind(exercise = df_1$exercise, exercise_F = df_1$exercise_F, sedentary = df_1$sedentary, sedentary_F = df_1$sedentary_F)


#strenuous_act:
#walking:
#sedentary:

```


## filter only non-NA data and output to a CSV file
```{r echo=TRUE}

COVID_na_rm <- COVID_cleaned %>% drop_na(PSQI_total,PSQI_F_total,DASS_total, DASS_F_total)

glimpse(COVID_na_rm)

write.csv(COVID_na_rm, "Score_Demographic_cleaned.csv",row.names = F, fileEncoding = "UTF-8")


```



## append descriptive questions
```{r}

descriptive_ <- df_1 %>% select(starts_with("add"),starts_with("seek"),starts_with("media"),starts_with("COVID"),starts_with("social"),starts_with("alcohol"),starts_with("smoke"),ends_with("impact"),starts_with("phyhealth"))

length(colnames(descriptive_))
?length

COVID_cleaned <- COVID_cleaned %>% cbind(descriptive_[1:length(colnames(descriptive_))])

COVID_na_rm <- COVID_cleaned %>% drop_na(PSQI_total,PSQI_F_total,DASS_total, DASS_F_total)

write.csv(COVID_cleaned, "COVID_cleaned.csv",row.names = F, fileEncoding = "UTF-8")

save.image(file='COVID_cleaned.RData')

```






## Codebook
see another pdf file for detailed description and corresponding variable value




Gender: 1-male, 2-female


Age, "Ethnicity", "Marital", "Education", "Occupation", "income", "household_size"

Eli1: Hong Kong Residence (non) permanent

Eli2: stayed in Hong Kong for most of the time in the past year

PSQI：
Q1 night shift -> exclude for irregular sleep
C4: bed time = (wakeup-gotosleep) + wake time? ambiguous wording? not add awake time in the doc I found
C5: Kris - 5.1 +..+ 5.8 , doc - 5.2 +..+ 5.9 + else
attrition: 1- 360/606

ISI: sum of the 7 items;

Seek:
show distribution for different choices

media: media consumption about COVID-19 and social movements
remain unttached

COVID:
form a score

Social: 


DASS: 
as dependent variable & independent
look at sub-scale too
(adjusted P value)
same as HABEAT, but score goes down in follow up survey
attrition: 1- 483/933

QoL: sum of the 5 items;

weight:
weight_2_1:
1 - Pounds
2 - kg
1 lbs = 0.45359237 kg
1 kg = 2.20462262 lbs

height:
height_2_1:
1 - cm
2 - feet/inches
1 feet = 30.48 cm = 12 inches
1 inch = 2.54 cm

（physical activity：）

strenuous_act:
walking:
sedentary:

alcohol:
1 - frequency
2 - fre binge drinking 

smoke:

social_impact:

virus_impact:

physical_health:

source:

---------follow up-----

PSQI: question listing slightly shifted from 1st survey

ISI: ''

DASS: ''

QoL: ''

strenuous_act:

walking:

sedentary:

Weight_F:

alcohol:

mayMarch/last7d

binge-drinking

retrospection
trend? pattern? 
correlation

media:

interestonlineinterv_F:

SC1:

ISI:


-----------------------


composite 
- media
- COVID
- phisical activity (multiply? add?)
- lifestyle (alcohol, smoking)

correlation 

regression 
(DASS, controling for baseline)
